---
title: "Exploring gapminder"
author: "Your Name"
output: html_document
---


```{r setup, include=FALSE}
# more substantial setup for your document
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      fig.width = 4, 
                      fig.height = 4, 
                      dev = "png",
                      cache = FALSE)
```

```{r library, echo = FALSE, message = FALSE}
library(tidyverse)
library(broom)
library(knitr)
```

````{r functions, echo = FALSE}
# A function to scale input to 0-1
scale_01 <- function(x){
  (x - min(x, na.rm = TRUE)) / diff(range(x, na.rm = TRUE))
}
```

```{r data-read-in}
gapminder <- read_csv(here::here("data","gapminder.csv"))
```

Here are the first 6 rows of data

```{r gapminder-top}
 top_gap <- head(gapminder, n = 10)
 bot_gap <- tail(gapminder, n = 10)
 
 knitr::kable(top_gap,
              caption = "The first 6 rows of the dataset, gapminder",
              digits = 2)
 kable(bot_gap)
```


```{r demo, eval = TRUE, echo = TRUE}
1 + 1
```

# Introduction

let's look at the lifespan

```{r hist-life-exp}
hist(gapminder$lifeExp)
```

Let's look at the GDP per capita

```{r example-cache}
hist(gapminder$gdpPercap)
```

Let's fit a simple linear model of the effect of year on life expectancy

```{r example-lm}
fit <- lm(lifeExp ~ year, data = gapminder)
fit
```

And let's look at the coefficient table:

```{r coef-table}
fit_coef <- tidy(fit)
knitr::kable(fit_coef,
             caption = "A table of the coefficients")
year_slope <- round(fit_coef$estimate[2], 3)
```

The effect of year on life expectancy is `r year_slope`.

# A table of the coefficients is:

```{r table-coef}
knitr::kable(fit_coef)
```

```{r}
```

```{r}

```

```{r}
1+1 

x <- 1 + 1 + 6
```

```{r gg-oz-plot, fig.cap = "Life expectancy from 1952 - 2007 for Australia. Life expentancy increases steadily except from 1962 to 1969. We can safely say that our life expectancy is higher than it has ever been!"}
library(ggplot2)
library(dplyr)
gapminder %>%
  filter(country == "Australia") %>%
  ggplot(aes(x = year,
             y = lifeExp)) + 
  geom_point()
```

Australia's life expectancy has increased a great deal over the past 50 years

(See Figure \@ref(fig:gg-oz-plot))


# References

